---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file - rmarkdown::render('README.Rmd', output_format = 'github_document', output_file = 'README.md') -->

```{r readme-setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/"
)
is_online = curl::has_internet()
```

# Geocomputation with R: Royal Geographic Society workshop 

[![Launch Rstudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/geocompr/RGSWorkshop/master?urlpath=rstudio)
[![Build Status](https://travis-ci.org/Robinlovelace/geocompr.svg?branch=master)](https://travis-ci.org/Robinlovelace/geocompr) [![](https://img.shields.io/docker/automated/robinlovelace/geocompr.svg)](https://hub.docker.com/r/robinlovelace/geocompr/builds/)

This repo contains a workbook to work through the [Geocomputation with R pre-conference workshop](https://www.rgs.org/events/summer-2019/geocomputation-with-r-%E2%80%93-free-workshop-and-reproduc/), an event sponsored by the [QMRG](https://quantile.info/geocomputation-with-r-book-demo-and-reproducible-map-competition/).

It is based on Geocomputation with R, a book by [Robin Lovelace](https://www.robinlovelace.net/), [Jakub Nowosad](https://nowosad.github.io/), and [Jannes Muenchow](http://www.geographie.uni-jena.de/en/Muenchow.html):

> Lovelace, Robin, Jakub Nowosad and Jannes Muenchow (2019). Geocomputation with R. The R Series. CRC Press.

This book has been published by [CRC Press](https://www.crcpress.com/9781138304512) in the [R Series](https://www.crcpress.com/Chapman--HallCRC-The-R-Series/book-series/CRCTHERSER). 
The online version of this book is free to read at https://geocompr.robinlovelace.net/.

## Getting started

This tutorial assumes you have installed a recent version of R, an editor you're comfortable with (we recommend RStudio), and the necessary packages.
If you have not, check out the guidance here: https://geocompr.robinlovelace.net/spatial-class.html#prerequisites
The packages we'll can be installed as follows^[
If you're running Mac or Linux, the previous command to install **sf** may not work first time.
These operating systems (OSs) have 'systems requirements' that are described in the package's [README](https://github.com/r-spatial/sf).
Various OS-specific instructions can be found online, such as the article *Installation of R 3.5 on Ubuntu 18.04* on the blog [rtask.thinkr.fr](https://rtask.thinkr.fr/blog/installation-of-r-3-5-on-ubuntu-18-04-lts-and-tips-for-spatial-packages/).
]:

```{r 02-spatial-data-1, eval=FALSE}
install.packages("sf")
install.packages("spData")
```

All the packages needed to reproduce the contents of the book can be installed with the following command: `devtools::install_github("geocompr/geocompkg")`.
The necessary packages can be 'loaded' (technically they are attached) with the `library()` function as follows:

```{r 02-spatial-data-3, message=FALSE}
library(sf)          # classes and functions for vector data
library(spData)      # contains example data
```

## What is geographic data in R?

In R geographic data is just an object like any other.
R's basic data analysis object type is the data frame, which looks like this:

```{r}
world
```

To plot geographic data, do:

```{r world-all, fig.cap="A spatial plot of the world using the sf package, with a facet for each attribute.", warning=FALSE, fig.scap="A spatial plot of the world using the sf package."}
plot(world)
```

